Lightness constancy under fixed geometry

### 2017-12-18 Draft1/draft1.tex
Worked on intro

The computational work reviewed above is based on formulating an explicit model of the image formation process and the statistical properties of natural images, and leveraging these formulations to obtain an appropriate "inverse optics" algorithm.  An alternative approach, which has been less-well explored, is to use supervised machine learning techniques to develop mappings between input images and stable color descriptors.  Along these lines, Barron (REF, that color balancing deep learning paper) recently developed a color balancing algorithm based on training a neural network ... (better go back and re-understand that paper so we can describe it here).  The use of supervised learning to understand human perceptual capabilities has enjoyed success in domains outside of color vision.  Geisler and colleagues, for example, have developed a technique they refer to as accuracy maximization analysis (AMA, REFS), which learns linear filters whose output carries information that is optimized for particular perceptual tasks. This technique has provided computations that efficiently estimate [depth, focus, whatever]  from image patches taken from natural scenes, and that in addition provide excellent models of human's ability to make corresponding estimates from the same stimuli.  In this paper, we apply AMA to a special case of the color constancy problem (lightness constancy) and evaluate how well it performs for naturalistic image input.

@@

Because of the difficulty in obtaining large labelled data sets of natural images for use in developing and evaluating computational color constancy algorithms, in this paper we adopt the approach of using high-quality computer graphics to render images of what we refer to as \em{virtual natural scenes}. We describe this approach in detail below, but the key advantage is that we can produce large numbers of naturalistic images where all of the underlying properties are well-specified. This allows us to explore and evaluate computational color constancy with rich stimuli, while retaining the ability to precisely control and manipulate the intrinsic properties of the objects and illuminants that comprise the scene. This general approach has been adopted by other labs for the study of computational estimation of optical flow (REF, Black) and is becoming increasingly popular (OTHER REFS: Cornell graphics person, Gallant, Yamins, Autonomous vehicle work), as well as for the generation of stimuli for the study of human perception (REFS: Many -- go to Boynton lecture slides for some).  Of course, success within the domain of graphics images does not guarantee smooth generalization to real natural images, a point we return to in the discussion.

In this initial work, we do not tackle the full problem of color constancy.  Rather, as a point of departure, we consider the computational problem of estimating a single attribute of object surface reflectance, which we refer to as surface luminance.  We compute object luminance from an object's surface reflectance function as follows.  We compute the light that would be reflected to the eye from a flat matte object with that surface reflectance, when the object is diffusely illuminated by the standard CIE daylight spectrum D65 (REF, CIE).  We then compute the luminance of this reflected light, using the CIE 1931 photopic luminosity function (REF, CIE), and divide this by the luminance of a perfectly reflecting diffuser (surface reflectance of one at all wavelengths) when that is computed in the same way.
%@techreport{RN1912,
%   author = {CIE},
%   title = {Colorimetry, second edition},
%   institution = {Bureau Central de la CIE},
%   number = {15.2},
%   year = {1986},
%   type = {Report}
%}
This procedure allows us to assign to any surface reflectance a single surface luminance value.  Because the perceived lightness of an object is approximately (but by no means exactly, see REF Wyszecki) correlated with surface luminance across a range of viewing conditions, we refer to the computational problem we study as that of lightness constancy.
%@inbook{RN10709,
%   author = {Wyszecki, G.},
%   title = {Color appearance},
%   booktitle = {Handbook of Perception and Human Performance: Sensory Processes and Perception},
%   publisher = {John Wiley & Sons},
%   address = {New York},
%   pages = {9.1-9.56},
%   year = {1986},
%   type = {Book Section}
%}
This is a generalization of the way this term is often used, where it refers more specifically to the stability of perceived lightness across images of scenes where all objects and illuminants are achromatic (REF, GILCHRIST BOOK).
%@book{RN3427,
%   author = {Gilchrist, A. L.},
%   title = {Seeing Black and White},
%   publisher = {Oxford University Press},
%   address = {Oxford},
%   year = {2006},
%   type = {Book}
%}
Here, we aim to estimate surface luminance, but do so for images where both the luminance and the chromaticity of the surfaces and illuminants varies (se  Fig.~\ref{fig:introExampleFigure}). A particularly important feature of our treatment is that we consider scene variations where not only the spectrum of the illumination varies, but also the surface reflectance of objects in the scene other than the object whose surface luminance is being estimeted.

The rest of this paper is organized as follows. [WRITE AN OUTLINE HERE.]  We conclude that [GIVE THE RESULT].

@@

\begin{comment}
\begin{figure}
\centering
\includegraphics[width=\textwidth]{approach.png}
\caption{{\bf Approach:} We have employed a supervised learning approach to study color constancy. We first generate a large dataset of multispectral images labeled with the lightness level of the target object. We generate images at multiple lightness levels, with images at the same lightness level having variable object reflectance spectra and illumination spectra. These multispectral images are used to simulate the response of the visual system. We use isetbio toolbox to model the optics and physiology of early vision and produce the response of the cones in the retina. We use supervised computational learning methods on this labeled dataset of cone-responses to estimate the lightness of the target object. The performance of the supervised method is calculated using a separate test dataset.} 
\label{fig:approach}
\end{figure}
\end{comment}

We employ a supervised learning approach to understand this phenomenon. We generate a database of multispectral images of naturalistic scenes, labeled by the lightness of a specific target object in the scene. These images are labeled with the perceived lightness of the target object under standard daylight spectrum $D_{65}$ (see SI). We generate hundreds of images with variations in the surface reflectance spectra of the objects and the illumination spectra of the light sources (Fig.~\ref{fig:introExampleFigure}). We study three cases resulting from the combination of such spectral variations (Fig.~\ref{fig:studiedCases}). These cases are: 1. Variation in target object spectrum (Fig.\ref{fig:targetVarying}) with fixed illumination and background, 2. Variations in both the target object spectrum and the light source spectra, with fixed background spectra (Fig.\ref{fig:targetIlluminantVarying}), and 3. Variations in all three spectra (target, background and light source) (Fig.\ref{fig:allSpectraVarying}). The effects of other types of spectral variations can be explained in terms of one of these cases (see Discussion/Supplementary information). Next, we simulate the response of the cones in the retina to these multispectral images using an accurate model of the early visual system. Such labeled cone-responses datasets are then analyzed using supervised computational learning methods. We use three methods: linear regression, generalized SVD regression and accuracy maximization analysis (AMA) in our analysis. These methods identify the computations that need to be performed on the cone-responses to estimate the target object lightness. These methods identify the receptive fields (RFs) that optimal for lightness estimation. We evaluate the performance of the learnt RFs on a test set of images.

% Figure 2: Cases Studied
\begin{figure}
\centering
	\begin{subfigure}[b]{0.33 \textwidth}
		\includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_a.jpeg}
		\caption{Case 1}
 		\label{fig:targetVarying}
	\end{subfigure}
	\begin{subfigure}[b]{0.33 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_b.jpeg}
        \caption{Case 2}
        \label{fig:targetIlluminantVarying}
    \end{subfigure}
	\begin{subfigure}[b]{0.33 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_c.jpeg}
        \caption{Case 3}
        \label{fig:allSpectraVarying}
    \end{subfigure}    
    \caption{{\bf Three types of spectral variations studied in this work:} sRGB rendition of multispectral images similar to ones studied in our work. The numbers on the left indicate the standard lightness level of the target object. 5 images are shown at each lightness level. We studied three types of spectral variations. (a) Case 1: Target reflectance spectrum variable, illumination spectra fixed, background spectra fixed, (b) Case 2: Target spectrum variable, Illumination spectra variable, Background spectra fixed, (c) Case 3: Target spectrum variable, Illumination spectra variable, Background spectra variable.}
\label{fig:studiedCases}
\end{figure}

In the following, we describe each one of these stages of our approach in detail. In the methods section, we first describe the process to generate naturalistic labeled images (\nameref{method:VirtualWorld}). Next, we describe the steps to simulate the response of the retinal cones to these images (\nameref{method:Isetbio}). Finally, we describe the three learning methods we have used in our analysis (\nameref{method:SupervisedLearning}). In the results section, we show the estimates and performance of each one of the supervised methods for the three cases with different spectral variations.

We show that for cases that have spectral variation only in target reflectance or background reflectance or the illuminant, the standard lightness ($Y_{65}$) can be estimated by simple transformations of the cone responses. For variations only in target object reflectance spectra (case 1), object lightness can be recovered directly from the light reflected off the target object. For case 2, where both the target object reflectance and light source spectra vary, object lightness can be recovered from the contrast between the target and the background. When all the spectral features in the scene are allowed to change (case 3), the standard lightness can not be recovered simply from the cone response or the contrast. Here, we have used two supervised learning methods with Bayesian estimators to estimate standard lightness of the target object. Using these methods, we can recover the lightness of the target object with $\sim 15\%$ relative root mean square error.

Our analysis shows that the optimal receptive fields for estimating the lightness of the target object show a center surround structure, supporting a comparison between the light coming from the target object and its surround. Additionally, lightness estimation involves a weighted sum of the L, M and S cone responses where the receptive fields give more weight to the L and M cone responses compared to S cone responses. Comparing the performance of the RFs learnt using  one case on the cone response of other cases shows that the receptive field of the complex cases perform well on images of simpler cases. We discuss the implications of such receptive fields on lightness constancy.

>>>>

The computational work reviewed above is based on formulating an explicit model of the image formation process and the statistical properties of natural images, and leveraging these formulations to obtain an appropriate "inverse optics" algorithm.  An alternative approach, which has been less-well explored, is to use supervised machine learning techniques to develop mappings between input images and stable color descriptors.  Along these lines, Barron (REF, that color balancing deep learning paper) recently developed a color balancing algorithm based on training a neural network ... (better go back and re-understand that paper so we can describe it here).  The use of supervised learning to understand human perceptual capabilities has enjoyed success in domains outside of color vision.  Geisler and colleagues, for example, have developed a technique they refer to as accuracy maximization analysis (AMA, REFS), which learns linear filters whose output carries information that is optimized for particular perceptual tasks. This technique has provided computations that efficiently estimate [depth, focus, whatever]  from image patches taken from natural scenes, and that in addition provide excellent models of human's ability to make corresponding estimates from the same stimuli.  In this paper, we apply AMA as well as comparison supervised learning algorithms to a special case of the color constancy problem (lightness constancy) and evaluate how well it performs for naturalistic image input. 
% [CHECK WHETHER ANY OF THIS NEEDS TO BE MELDED INTO THE PARAGRAPH ABOVE]
%We employ a supervised learning approach to understand this phenomenon. We generate a database of multispectral images of naturalistic scenes, labeled by the lightness of a specific target object in the scene. These images are labeled with the perceived lightness of the target object under standard daylight spectrum $D_{65}$ (see SI). We generate hundreds of images with variations in the surface reflectance spectra of the objects and the illumination spectra of the light sources (Fig.~\ref{fig:introExampleFigure}). We study three cases resulting from the combination of such spectral variations

@@

Because of the difficulty in obtaining large labelled data sets of natural images for use in developing and evaluating computational color constancy algorithms, in this paper we adopt the approach of using high-quality computer graphics to render images of what we refer to as {\em virtual natural scenes}. In addition to employing computer graphics to render the light reaching the eye from specified scenes, we also model the initial steps of the human visual system, so that our supervised learning methods are applied not to image pixels but rather to modeled responses of the foveal cone photoreceptor mosaic.  We describe this approach in detail below, but the key advantage is that we can produce large numbers of naturalistic images and corresponding cone mosaic responses, where all of the underlying scene properties are well-specified. This allows us to explore and evaluate computational color constancy with rich stimuli, while retaining the ability to precisely control and manipulate the intrinsic properties of the objects and illuminants that comprise the scene. This general approach has been adopted by other labs for the study of computational estimation of optical flow (REF, Black) and is becoming increasingly popular (OTHER REFS: Cornell graphics person, Gallant, Yamins, Autonomous vehicle work), as well as for the generation of stimuli for the study of human perception (REFS: Many -- go to Boynton lecture slides for some).  Of course, success within the domain of graphics images does not guarantee smooth generalization to real natural images, a point we return to in the \ref{Discussion}.

In this initial work, we do not tackle the full problem of color constancy.  Rather, as a point of departure, we consider the computational problem of estimating a single attribute of object surface reflectance, which we refer to as surface luminance.  We compute object luminance from an object's surface reflectance function as follows.  We compute the light that would be reflected to the eye from a flat matte object with that surface reflectance, when the object is diffusely illuminated by the standard CIE daylight spectrum D65 (REF, CIE).  We then compute the luminance of this reflected light, using the CIE 1931 photopic luminosity function (REF, CIE), and divide this by the luminance of a perfectly reflecting diffuser (surface reflectance of one at all wavelengths) when that is computed in the same way.
%@techreport{RN1912,
%   author = {CIE},
%   title = {Colorimetry, second edition},
%   institution = {Bureau Central de la CIE},
%   number = {15.2},
%   year = {1986},
%   type = {Report}
%}
This procedure allows us to assign to any surface reflectance a single surface luminance value.  Because the perceived lightness of an object is approximately (but by no means exactly, see REF Wyszecki) correlated with surface luminance across a range of viewing conditions, we refer to the computational problem we study as that of {\em lightness constancy}.
%@inbook{RN10709,
%   author = {Wyszecki, G.},
%   title = {Color appearance},
%   booktitle = {Handbook of Perception and Human Performance: Sensory Processes and Perception},
%   publisher = {John Wiley & Sons},
%   address = {New York},
%   pages = {9.1-9.56},
%   year = {1986},
%   type = {Book Section}
%}
This is a generalization of the way this term is often used, where it refers more specifically to the stability of perceived lightness across images of scenes where all objects and illuminants are achromatic (REF, GILCHRIST BOOK).
%@book{RN3427,
%   author = {Gilchrist, A. L.},
%   title = {Seeing Black and White},
%   publisher = {Oxford University Press},
%   address = {Oxford},
%   year = {2006},
%   type = {Book}
%}
Here, we aim to estimate surface luminance, but do so for images where both the luminance and the chromaticity of the surfaces and illuminants varies (see  Fig.~\ref{fig:introExampleFigure}). A particularly important feature of our treatment is that we consider scene variations where not only the spectrum of the illumination varies, but also the surface reflectance of objects in the scene other than the object whose surface luminance is being estimated.

Even within the special case of lightness constancy, there are many scene manipulations that could be considered. The set of manipulations we focus on here are illustrated by the three panels of (Fig.~\ref{fig:studiedCases}). These cases are: 1. variation in target object spectrum (Fig.\ref{fig:targetVarying}) with fixed light source and background objects spectra, 2. variations in both the target object spectrum and the light source spectra, with fixed background object spectra (Fig.\ref{fig:targetIlluminantVarying}), and 3. variations in all three types of spectra (target object, light source, and background objects) (Fig.\ref{fig:allSpectraVarying}). Across real scenes, there will also be variations in the shape of the target object, its pose, eye position, etc. (see \ref{Methods}). In this paper, we focus on the problem of lightness constancy within scene ensembles where these geometric factors are held fixed.

The rest of this paper is organized as follows. [WRITE AN OUTLINE HERE.]  We conclude that [GIVE THE RESULT].
%EDIT THIS IN HERE.
%In the following, we describe each one of these stages of our approach in detail. In the methods section, we first describe the process to generate naturalistic labeled images (\nameref{method:VirtualWorld}). Next, we describe the steps to simulate the response of the retinal cones to these images (\nameref{method:Isetbio}). Finally, we describe the three learning methods we have used in our analysis (\nameref{method:SupervisedLearning}). In the results section, we show the estimates and performance of each one of the supervised methods for the three cases with different spectral variations. 
%
%We show that for cases that have spectral variation only in target reflectance or background reflectance or the illuminant, the standard lightness ($Y_{65}$) can be estimated by simple transformations of the cone responses. For variations only in target object reflectance spectra (case 1), object lightness can be recovered directly from the light reflected off the target object. For case 2, where both the target object reflectance and light source spectra vary, object lightness can be recovered from the contrast between the target and the background. When all the spectral features in the scene are allowed to change (case 3), the standard lightness can not be recovered simply from the cone response or the contrast. Here, we have used two supervised learning methods with Bayesian estimators to estimate standard lightness of the target object. Using these methods, we can recover the lightness of the target object with $\sim 15\%$ relative root mean square error. 
%
%Our analysis shows that the optimal receptive fields for estimating the lightness of the target object show a center surround structure, supporting a comparison between the light coming from the target object and its surround. Additionally, lightness estimation involves a weighted sum of the L, M and S cone responses where the receptive fields give more weight to the L and M cone responses compared to S cone responses. Comparing the performance of the RFs learnt using  one case on the cone response of other cases shows that the receptive field of the complex cases perform well on images of simpler cases. We discuss the implications of such receptive fields on lightness constancy.

@@

% Figure 2: Cases Studied
\begin{figure}
\centering
	\begin{subfigure}[b]{0.33 \textwidth}
		\includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_a.jpeg}
		\caption{Case 1}
 		\label{fig:targetVarying}
	\end{subfigure}
	\begin{subfigure}[b]{0.33 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_b.jpeg}
        \caption{Case 2}
        \label{fig:targetIlluminantVarying}
    \end{subfigure}
	\begin{subfigure}[b]{0.33 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_c.jpeg}
        \caption{Case 3}
        \label{fig:allSpectraVarying}
    \end{subfigure}    
    \caption{{\bf Three types of spectral variations studied in this work:} sRGB rendition of multispectral images similar to ones studied in our work. The numbers on the left indicate the standard lightness level of the target object. 5 images are shown at each lightness level. We studied three types of spectral variations. (a) Case 1: target object relative surface reflectance spectrum variable, light source spectra fixed, background object spectra fixed, (b) Case 2: target object relative surface reflectance spectrum variable, light source spectra variable, background object spectra fixed, (c) Case 3: target object relative surface reflectance spectrum variable, light source spectra variable, background object spectra variable. This panel is reproduced from \ref{fig:introExampleFigure}. As in \ref{fig:introExampleFigure}, the spheres in each row of each panel have the same surface luminance, while the spheres in each column of each panel have the same relative surface reflectance.  Moreover, across the three panels of this figure, spheres in corresponding locations have the same surface reflectance. In all three panels, the overall lights source spectra scale factors (see \ref{Methods}) were drawn from a uniform distribution on the range [0.5, 1.5]. [CHECK THAT THIS WORKS FOR A NICE FIGURE].  The sRGB values for all three panels were normalized using a common scale factor prior to gamma correction.} 
\label{fig:studiedCases}
\end{figure}
